---
title: "Statistical Inference Project Part 1"

author: "Damien Edwards"
date: "January 19, 2015"
output: pdf_document
---
This project will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution function is rexp(n,lambda) and lambda is the rate parameter.The mean of exponential distribution is 1/lambda and the standard deviation is also       1/lambda. Set lambda = 0.2 for all of the simulations.

The simulation shown below in R used to calculate the mean:
```{r}
lambda = 0.2
n = 40
nsims = 1000
set.seed(240)
## Theoretical values
theo_mean <- 1/lambda
theo_sigma <- 1/lambda /sqrt(n)
```

```{r}
simulate_matrix <- matrix(rexp(nsims * n, rate=lambda), nsims, n)
simulate_mean <- rowMeans(simulate_matrix)
```


The simulation matrix data is shown below for expoloration

```{r, echo=FALSE}
hist(simulate_mean, col = "blue", main="Histogram of exponential distribution")
```

##Comparison of Sample Mean and Theoretical Mean
```{r echo=TRUE}
compare_mean<-mean(simulate_mean)
theo_mean <- 1/lambda

```
```{r echo=TRUE}
## Mean value to compare
compare_mean
## Theoretical mean value
theo_mean
```
The distribution based on the simulations is 5.026972 and the theoretical mean value is 5. The means are close to the same center of distribution.


## Comparison of Variance for the sample data and Theoretical variance
```{r echo=TRUE}
actual_var <- var(simulate_mean)
theo_var <- (1/lambda)^2/n
## Actual variance
actual_var

## Theoretical variance
theo_var
```
The actual variance value is .599 and the theoretical variance is .625. The values are close.

## The distribution is approximately normal
The histogram plot of the means of the 1000 simulations of rexp(n, lambda) shown below.

```{r, echo=FALSE}
library(ggplot2)
df <- data.frame(simulate_mean);
plot_1 <- ggplot(df, aes(x =simulate_mean))
plot_1 <- plot_1 + geom_histogram(aes(y=..density..), colour="black",
fill = "green")
plot_1 + geom_density(colour="blue", size=1)+  xlab("means of observation");
```
##
The variance and mean of the sample data are similar and resemble a normal distribution. Below is the confidence intervals calculated:
```{r, echo=TRUE}
actual_interval <- round (mean(simulate_mean) + c(-1,1)*1.96*sd(simulate_mean)/sqrt(n),3)
theory_interval <- theo_mean + c(-1,1)*1.96*sqrt(theo_var)/sqrt(n);

##Actual confidence interval
actual_interval
## Theoretical confidence interval
theory_interval
```
Actual 95% confidence interval [4.787, 5.267]. Theoretical 95% confidence interval [4.755, 5.245]
This shows the distribution is approximately normal.



